Competing Bandits:\\ The Perils of Exploration under Competition \thanks{This is an extended abstract of \citet{CompetingBandits-merged}, which in turn subsumes conference papers \citet{CompetingBandits-itcs18} and \citet{CompetingBandits-ec19}.}

### 2020-09-22 journal paper/content/sec-model.tex
Lots of edits re equilibria (mixed vs pure, finite vs infinite, etc.)

\xhdr{Competition game.}
Some of our results explicitly study the game between the two principals, termed the \emph{competition game}. We model it as a simultaneous-move game: before the first agent arrives, each principal commits to an MAB algorithm.
\asedit{Pure strategies in this game correspond to MAB algorithms.}%
\footnote{\asedit{Note that a mixed strategy is simply a randomized MAB algorithm, \ie it is also a pure strategy.}}
 Principal's utility is defined as the market share, \ie the number of agents that chose this principal. Principals are risk-neutral: they optimize their expected utility.

>>>>

\xhdr{Competition game.}
Some of our results explicitly study the game between the two principals, termed the \emph{competition game}. We model it as a simultaneous-move game: before the first agent arrives, each principal commits to an MAB algorithm.
\asedit{Pure strategies in this game correspond to MAB algorithms.}%
\footnote{\asedit{Note that a mixed strategy is also a (randomized) MAB algorithm, \ie also a pure strategy.}}
 Principal's utility is defined as the market share, \ie the number of agents that chose this principal. Principals are risk-neutral: they optimize their expected utility.

### 2020-09-22 journal paper/content/sec-theory.tex
Lots of edits re equilibria (mixed vs pure, finite vs infinite, etc.)

We posit fully rational agents, in the sense that their response function is \HardMax. We show that principals are not incentivized to \emph{explore}--- \ie to deviate from \DynGreedy. The core technical result is that if one principal adopts \DynGreedy, then the other principal loses all agents as soon as he deviates.

To make this more precise, let us define what it means for MAB algorithms to deviate from each other.

@@

\begin{corollary}\label{cor:DG-dominance}
The competition game has a unique Nash equilibrium: both principals choose \DynGreedy. This holds even if the principals can only choose from restricted strategy sets which include \DynGreedy.
\end{corollary}

The proof of Theorem~\ref{thm:DG-dominance} relies on two key lemmas: that deviating from \DynGreedy implies a strictly smaller Bayesian-expected reward, and that \HardMax implies a ``sudden-death" property: if one agent chooses principal $1$ with certainty, then so do all subsequent agents. We re-use both lemmas in later results, so we state them in sufficient generality. In particular, Lemma~\ref{lm:DG-rew} works for any response function because it only considers the performance of each algorithm without competition.

\begin{lemma}\label{lm:DG-rew}
Assume that \alg[1] is \DynGreedy, and \alg[2] deviates from \DynGreedy starting from some (local) step $n_0<T$. Then $\rew_1(n_0)>\rew_2(n_0)$. This holds for any response function $\respF$.
\end{lemma}

@@

\begin{theorem}\label{thm:random-clean}
Assume \HardMaxRandom response function. Suppose algorithms \alg[1], \alg[2] are \bmonotone and $T$-oblivious, and \eqref{eq:random-assn} holds. If \alg[1] \emph{BIR-dominates} \alg[2] starting from step $n_0$, then each agent $t\geq \max(n_0,m_0)$ chooses principal $1$ with probability $\respF(1) = 1- \eps_0$ (which is the largest possible).
\end{theorem}

@@

We'd like to use Theorem~\ref{thm:random-clean} to conclude that a (much) better algorithm prevails in equilibrium. Indeed, this holds if the principals can only choose from a finite set of MAB algorithms.

\begin{definition}\label{def:restricted-competition}
A \emph{\FiniteGame} is the competition game between the two principals in which they can only choose from a finite set of MAB algorithms (called \emph{feasible}). All feasible algorithms are assumed to be
\bmonotone and $T$-oblivious, and to satisfy \eqref{eq:random-assn}.
\end{definition}

\begin{corollary}\label{cor:random}
Assume \HardMaxRandom response function. Consider the \FiniteGame with a feasible algorithm \alg which BIR-dominates all other feasible algorithms. Then, for any sufficiently large time horizon $T$, this game has a unique Nash equilibrium, in which both principals choose algorithm \alg.
\end{corollary}

@@

We find that the greedy modification prevails in competition if $p$ is small enough.

@@

Let us show how \alg[1] can be a better algorithm than its greedy modification \alg[2]. We need to assume that \alg[1] is (much) better than the greedy step \alggr if the latter is computed on a fraction of the data. Formally, we posit a convex function $f(\cdot)$ and $m_0>0$ such that
\[ \BIRgr(n) \geq f(n)
    \quad\text{and}\quad
    f\rbr{\E[M_n]} \geq 2\, \BIR_1(n) \quad \forall n\geq m_0 .  \]
Then by \eqref{eq:BIR-modification} and Jensen's inequality, for each step $n\geq m_0$ we have
\[\BIR_2(n)
    \geq  \E\sbr{ \BIRgr(M_n) }
    \geq \E\sbr{ f(M_n) }
    \geq f\rbr{ \E[M_n] }
    \geq 2\,\BIR_1(n). \]

@@

\begin{corollary}\label{cor:random-greedy}
Posit a symmetric \HardMaxRandom response function. Suppose both principals can only choose from \bmonotone MAB algorithms. Then for any time horizon $T$, the only possible \emph{pure} Nash equilibrium is one where both principals choose \DynGreedy. Moreover, no pure Nash equilibrium exists when some algorithm BIR-dominates \DynGreedy and $T$ is sufficiently large.
These conclusions hold even if each principal's strategy set is a subset of \bmonotone MAB algorithms that is closed under the greedy modification.%
\footnote{That is, the strategy set of each principal consists of some (but not necessarily all) \bmonotone MAB algorithms, and contains the greedy modification for each algorithms in this set and each mixing parameter $p$.}
\end{corollary}

@@

As in the previous subsection, we can state a game-theoretic corollary for the \FiniteGame (as in Definition~\ref{def:restricted-competition}). We need a mild technical assumption that cumulative Bayesian regret (\BReg) tends to infinity. \BReg is a standard notion from the literature (along with \BIR):
\begin{align}\label{eq:SoftMax-BReg}
\BReg(n) := n\cdot \E_{\mu\sim\priorMu}
    \left[ \max_{a\in A} \mu_a\right] -
    \E\sbr{ \sum_{m=1}^n \rew(m)}
    = \sum_{m=1}^n \BIR(m).
\end{align}

\begin{corollary}\label{cor:SoftMax-strong}
Assume \SoftMaxRandom response function. Consider the \FiniteGame in which all feasible algorithms satisfy $\BReg(n)\to \infty$.%
\footnote{$\BReg(n)\to \infty$ is a mild non-degeneracy condition, see Appendix~\ref{app:bg} for background.}
Suppose a feasible algorithm \alg BIR-dominates all others. Then, for any sufficiently large time horizon $T$, there is a unique Nash equilibrium: both principals choose \alg.
\end{corollary}

>>>>

We posit fully rational agents, in the sense that their response function is \HardMax. We show that principals are not incentivized to \emph{explore}--- \ie to deviate from \DynGreedy. The core technical result is that if one principal adopts \DynGreedy, then the other principal loses all agents as soon as he deviates, in the sense defined below.

@@

\asedit{\DynGreedy is a weakly dominant strategy in the competition game, and a unique Nash equilibrium.  This is because \DynGreedy receives, in expectation, at least half of the agents before global round $n_0$, and all agents after that; both are the best possible against \alg[2]. Moreover, \DynGreedy guarantees at least $T/2$ agents in expectation, and any other strategy can receive strictly less than that (\eg if the opponent chooses \DynGreedy).}

\begin{corollary}\label{cor:DG-dominance}
\asedit{\DynGreedy is a weakly dominant strategy in the competition game.} The game has a unique Nash equilibrium: both principals choose \DynGreedy.
\end{corollary}

The proof of Theorem~\ref{thm:DG-dominance} relies on two key lemmas: that deviating from \DynGreedy implies a strictly smaller Bayesian-expected reward, and that \HardMax implies a ``sudden-death" property: if one agent chooses principal $1$ with certainty, so do all subsequent agents. We re-use both lemmas in later results, so we state them in sufficient generality.

%In particular, Lemma~\ref{lm:DG-rew} works for any response function, as it only considers the stand-alone performance of each algorithm.

\begin{lemma}\label{lm:DG-rew}
Assume that \alg[1] is \DynGreedy, and \alg[2] deviates from \DynGreedy starting from some (local) step $n_0<T$. Then $\rew_1(n_0)>\rew_2(n_0)$. The lemma holds for any response function $\respF$ (as it only considers the stand-alone performance of each algorithm).
\end{lemma}

@@

\begin{theorem}\label{thm:random-clean}
Fix a \HardMaxRandom response function $\respF$. Suppose algorithms \alg[1], \alg[2] are \bmonotone and $T$-oblivious, and \eqref{eq:random-assn} holds. If \alg[1] \emph{BIR-dominates} \alg[2] starting from step $n_0$, then each agent $t\geq \max(n_0,m_0)$ chooses principal $1$ with probability $\respF(1) = 1- \eps_0$ (which is the largest possible selection probability for this response function).
\end{theorem}

@@

We'd like to use Theorem~\ref{thm:random-clean} to conclude that a (much) better algorithm prevails in equilibrium. \asedit{We consider a version of the competition game in which the principals are restricted to choosing from a given set of MAB algorithms; the algorithms in this set are called \emph{feasible}.}

%Indeed, this holds if the principals can only choose from a finite set of MAB algorithms.
%\begin{definition}\label{def:restricted-competition}
%A \emph{\FiniteGame} is the competition game between the two principals in which they can only choose from a finite set of MAB algorithms (called \emph{feasible}). All feasible algorithms are assumed to be \bmonotone and $T$-oblivious, and to satisfy \eqref{eq:random-assn}.
%\end{definition}

\begin{corollary}\label{cor:random}
Fix a \HardMaxRandom response function $\respF$. \asedit{Consider the competition game in which all feasible MAB algorithms are $T$-oblivious, \bmonotone, and satisfy \eqref{eq:random-assn} for some fixed $m_0$. Suppose some feasible algorithm \alg  BIR-dominates all other feasible algorithms, starting from some local step $n_0$.} Then, for any sufficiently large time horizon $T$, this game has a unique Nash equilibrium, in which both principals choose algorithm \alg.
\end{corollary}

@@

We find that the greedy modification prevails in competition if $p$ is small enough. We focus on \emph{symmetric} response functions: ones that satisfy 
$\respF(x)+\respF(-x)=1$ for any $x\in[0,1]$.

@@

First, let us show how \alg[1] can be a better algorithm than its greedy modification \alg[2]. We need to assume that \alg[1] is (much) better than the greedy step \alggr if the latter is computed on a fraction of the data. Formally, we posit a convex function $f(\cdot)$ and $m_0>0$ such that
\[ \BIRgr(n) \geq f(n)
    \quad\text{and}\quad
    f\rbr{\E[M_n]} \geq 2\, \BIR_1(n) \quad \forall n\geq m_0 .  \]
Then by \eqref{eq:BIR-modification} and Jensen's inequality, for each step $n\geq m_0$ we have
\[\BIR_2(n)
    \geq  \E\sbr{ \BIRgr(M_n) }
    \geq \E\sbr{ f(M_n) }
    \geq f\rbr{ \E[M_n] }
    \geq 2\,\BIR_1(n). \]

@@

\begin{corollary}\label{cor:random-greedy}
Fix a symmetric \HardMaxRandom response function $\respF$. \asedit{Consider the competition game in which algorithms are feasible if and only if they are \bmonotone.}%
\footnote{\asedit{More generally, this corollary holds if feasible algorithms are \bmonotone and closed under the greedy modification (\ie the greedy modification of any feasible algorithm is also feasible, for any mixing parameter $p$).}}
\asedit{Then there are no Nash equilibria other than (\DynGreedy,\,\DynGreedy).}
%The latter is not a Nash equilibrium when some algorithm BIR-dominates \DynGreedy and time horizon $T$ is sufficiently large.
\end{corollary}

\asedit{Recall that by Theorem~\ref{thm:random-clean} \DynGreedy cannot be played in equilibrium when it is BIR-dominated by some feasible, $T$-oblivious algorithm and the time horizon $T$ is large enough.}

%These conclusions hold even if each principal's strategy set is a subset of \bmonotone MAB algorithms that is closed under the greedy modification.%
%\footnote{That is, the strategy set of each principal consists of some (but not necessarily all) \bmonotone MAB algorithms, and contains the greedy modification for each algorithms in this set and each mixing parameter $p$.}

@@

\asedit{It follows that a weakly-BIR-dominating algorithm prevails in equilibrium.} We need a mild technical assumption that cumulative Bayesian regret (\BReg) tends to infinity. \BReg is a standard notion from the literature (along with \BIR):
\begin{align}\label{eq:SoftMax-BReg}
\BReg(n) := n\cdot \E_{\mu\sim\priorMu}
    \left[ \max_{a\in A} \mu_a\right] -
    \E\sbr{ \sum_{m=1}^n \rew(m)}
    = \sum_{m=1}^n \BIR(m).
\end{align}

\begin{corollary}\label{cor:SoftMax-strong}
Assume \SoftMaxRandom response function. Consider the competition game in which all feasible algorithms are \bmonotone, $T$-oblivious, and satisfy $\BReg(n)\to \infty$.%
\footnote{$\BReg(n)\to \infty$ is a mild non-degeneracy condition, see Appendix~\ref{app:bg} for background.}
Suppose some feasible algorithm \alg weakly-BIR-dominates all others. Then, for any sufficiently large time horizon $T$, there is a unique Nash equilibrium: both principals choose \alg.
\end{corollary}
